---
title: "Final Project Report"
author: "Brian Anderson and Daniel Beidelschies"
output: 
  html_document:
    df_print: paged
    code_download: true
    theme: cerulean
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=TRUE, message=FALSE, warning=FALSE)
```

```{r libraries, echo = FALSE}
library(tidyverse)     # for graphing and data cleaning
library(lubridate)     # for date manipulation
library(ggthemes)      # for even more plotting theme
library(gganimate)     # for adding animation layers to ggplots
library(transformr)    # for "tweening" (gganimate)
library(gifski)        # for creating the gif
library(shiny)         # for creating interactive apps
library(readr)
library(patchwork)
theme_set(theme_minimal()) # My favorite ggplot() theme :)
```

## Data choice: Video Game Sales

```{r, echo = FALSE}
vgsales <- read_csv("vgsales.csv", col_types = cols(Year = col_integer() ))
vgsales
```


### Title: 

In the YAML section of the document. A descriptive title & list of all group members.

**Exploring the World of Video Games**

Daniel Beidelschies and Brian Anderson

### Introduction and Background: 

An introduction that motivates & outlines a clear, specific set of research questions. Also, provide some background on your topic.

##### Introduction

Since the advancement of technology, the creation of video games has played an influential role in being a source of entertainment for everyone around the world. From the multitude of video games that have been created, everyone is able to find something that they enjoy in order to pass time, have fun, and fulfill their creative needs. With this idea, we wanted to look at how well different games, genres, and platforms did when it came to releasing games for everyone to play. This will allow us to understand how popularity ranks up between different types of games based on the sales around of the world. In addition, we were also curious on how much revenue this industry makes as we spent our own money to fully experience these games. 

#### Background information

Since our individual discovery, we have always had interest in video games and wondered how the sales of these games differed across the world as we grew up in different regions. Furthermore, the video games industry has gained popularity and interactivity over the years with the release of evolving tech and we wanted to see how these sales were reflected. 

### Research Questions

What are the most selling genres of games over the years? And across different regions?

What are the most selling platforms of games over the years across different regions? Is there connection between platforms sold and genre popularity?

What are the most selling game franchises based on global sales?

What are the ten most popular games over the years?

### Data collection: 

Specification of your data sources and collection process.

Scraping from Kaggle and importing to a dataset

### Analysis: 

This is the bulk of the report which either has a presentation of the group’s key findings and take-aways or gives the detail of how someone would interact and what people should take away from the shiny app. If you choose to do a shiny app, be sure to include a link to the shinyapps.io site.

At first, we looked at which genre of games was the most popular based on global sales. 

```{r, fig.alt = "Animated Line Graph showing cumulative video game sales by genre from 1980-2020.", echo=FALSE}
vgsales %>%
  group_by( Year,Genre) %>% 
  summarize(total_genre_sum = sum(Global_Sales)) %>% 
  ungroup() %>% 
  complete(Year,Genre) %>% 
  arrange(Genre,Year) %>% 
  group_by(Genre) %>% 
  replace_na(list(total_genre_sum = 0)) %>% 
  mutate(cum_genre_sum = cumsum(total_genre_sum)) %>%
  filter(cum_genre_sum > 0) %>% 
  filter(n() >= 13) %>% 
  arrange(Year,desc(cum_genre_sum)) %>%
  drop_na() %>% 
  mutate(rank = 1:n()) %>% 
  ggplot(aes(x = Year,
             y = cum_genre_sum,
             color = Genre)) +
  geom_line(size = 1.5) +
  labs(title = "Popular Video Game Genres based on Worldwide Sales",
       subtitle = "Year: {frame_along}",
       x = "", 
       y = "",
       caption = "Plot created by D.B. & B.A., Data from Vgsales") +
  scale_color_manual(values = c("red", "green","blue","black", "yellow","deeppink","cyan","darkmagenta","steelblue","midnightblue","sienna", "green4"))+
  transition_reveal(Year)
```

From this animated graph, we saw that the Action genre was the most popular by the end of the time period in 2020. It was followed by Sports, Shooting, Role-playing, and Platform. At the beginning of the graph, platform was most selling until around the early 2000s when Action, Sports, Role-playing took a major incline. Furthermore, Action and experienced a gradual linear path whereas Shooting, Role-Playing, and Platform followed a steady linear path. 

Next, we looked at the total sales of each genre in different locations around the world including North America, Europe, Japan, and other areas that aren't considered from the former three. 

```{r, fig.alt = "Column chart showing the total global sales of each genre in different locations.", echo = FALSE}
regiongenresales <- vgsales %>%
  group_by( Genre) %>% 
  summarize(total_NA_genre_sum = sum(NA_Sales), 
            total_EU_genre_sum = sum(EU_Sales), 
            total_JP_genre_sum = sum(JP_Sales),
            total_other_genre_sum = sum(Other_Sales)) %>% 
  ungroup() %>% 
  complete(Genre) %>% 
  arrange(Genre) %>% 
  group_by(Genre) %>% 
  replace_na(list(total_NA_genre_sum = 0, 
                  total_EU_genre_sum = 0, 
                  total_JP_genre_sum = 0, 
                  total_other_genre_sum = 0)) %>% 
  filter(total_NA_genre_sum > 0)


g1 <- ggplot(regiongenresales) +
  geom_col(aes(x = total_NA_genre_sum,
               y = fct_reorder(Genre, total_NA_genre_sum),
               fill = factor(Genre))) +
  labs(x = "", 
       y = "", 
       caption = "Plot created by D.B. & B.A., Data from Vgsales",
       title = "North American Sales by Genre") +
  theme(legend.position = "none") +
  scale_fill_manual(values = c("Action" = "red", 
                               "Role-Playing" = "blue"))

g2 <- ggplot(regiongenresales) +
  geom_col(aes(x = total_EU_genre_sum,
               y = fct_reorder(Genre, total_EU_genre_sum),
               fill = factor(Genre))) +
  labs(x = "", 
       y = "",
       caption = "Plot created by D.B. & B.A., Data from Vgsales",
       title ="European Sales by Genre") +
  theme(legend.position = "none") +
  scale_fill_manual(values = c("Action" = "red", 
                               "Role-Playing" = "blue"))

g3 <- ggplot(regiongenresales) +
  geom_col(aes(x = total_JP_genre_sum,
               y = fct_reorder(Genre, total_JP_genre_sum),
               fill = factor(Genre))) +
  labs(x = "", 
       y = "",
       caption = "Plot created by D.B. & B.A., Data from Vgsales",
       title = "Japanese Sales by Genre") +
  theme(legend.position = "none") +
  scale_fill_manual(values = c("Action" = "red", 
                             "Role-Playing" = "blue"))

g4 <- ggplot(regiongenresales) +
  geom_col(aes(x = total_other_genre_sum,
               y = fct_reorder(Genre, total_other_genre_sum),
               fill = factor(Genre))) +
  labs(x = "", 
       y = "",
       caption = "Plot created by D.B. & B.A., Data from Vgsales",
       title = "Other Region Sales by Genre") +
  theme(legend.position = "none") +
  scale_fill_manual(values=c("Action" = "red", 
                             "Role-Playing"= "blue"))


g1
g2
g4
g3
```


From these graphs, we found that Action remained the highest selling genre in North America, Europe, and other regions. Although, it was a different story for Japan as Role-playing was the most popular genre, about two times higher than the second competing genre of Action. 

With this information, we then wanted to see how different genres were spread throughout the different platforms and determine if there was a connection between the platforms and the sales of specific genres. 


```{r, fig.height = 10, fig.width = 12, fig.alt = "Column chart showing the total sales of each platform in different locations with the distribution of each genre.", echo = FALSE}
regionplatformsales <- vgsales %>%
  group_by(Platform, Genre) %>% 
  summarize(total_NA_platform_sum = sum(NA_Sales), 
            total_EU_platform_sum = sum(EU_Sales), 
            total_JP_platform_sum = sum(JP_Sales),
            total_other_platform_sum = sum(Other_Sales),
            total_global_platform_sum = sum(Global_Sales)) %>% 
  ungroup() %>% 
  complete(Platform) %>% 
  arrange(Platform) %>% 
  group_by(Platform) %>% 
  replace_na(list(total_NA_platform_sum = 0, total_EU_platform_sum  = 0, total_JP_platform_sum  = 0, total_other_platform_sum = 0)) %>% 
  mutate(cum_NA_plat_sum = cumsum(total_NA_platform_sum), 
         cum_EU_plat_sum = cumsum(total_EU_platform_sum),
         cum_JP_plat_sum = cumsum(total_JP_platform_sum),
         cum_other_plat_sum = cumsum(total_other_platform_sum),
         cum_global_plat_sum = cumsum(total_global_platform_sum)) 
  
p1 <- regionplatformsales %>% 
  filter(total_NA_platform_sum >0) %>% 
  ggplot()+
  geom_col(aes(x = total_NA_platform_sum,
             y = fct_reorder(Platform,cum_NA_plat_sum),
             fill = factor(Genre),
             group = Platform))+
  scale_fill_manual(values = c("red", "green","blue","black", "yellowgreen", "deeppink","cyan","darkmagenta","steelblue","midnightblue","gold", "green4"))+
  labs(x = "", 
       y = "",
       fill = "Genre",
       caption = "Plot created by D.B. & B.A., Data from Vgsales",
       title = "North American Sales by Platform")

p2 <- regionplatformsales %>% 
  replace_na(list(cum_global_plat_sum = 0)) %>%
  filter(cum_global_plat_sum >0) %>% 
  ggplot()+
  geom_col(aes(x = total_EU_platform_sum,
             y = fct_reorder(Platform,cum_EU_plat_sum),
             fill = factor(Genre),
             group = Platform))+
  scale_fill_manual(values = c("red", "green","blue","black", "yellowgreen", "deeppink","cyan","darkmagenta","steelblue","midnightblue","gold", "green4")) +
  labs(x = "", 
       y = "",
       fill = "Genre",
       caption = "Plot created by D.B. & B.A., Data from Vgsales",
       title = "European Sales by Platform")

p3 <- regionplatformsales %>% 
  filter(total_JP_platform_sum >0) %>% 
  ggplot()+
  geom_col(aes(x = total_JP_platform_sum,
             y = fct_reorder(Platform,cum_JP_plat_sum),
             fill = factor(Genre),
             group = Platform))+
  scale_fill_manual(values = c("red", "green","blue","black", "yellowgreen", "deeppink","cyan","darkmagenta","steelblue","midnightblue","gold", "green4"))+
  labs(x = "", 
       y = "",
       fill = "Genre",
       caption = "Plot created by D.B. & B.A., Data from Vgsales",
       title ="Japanese Sales by Platform")

p4 <- regionplatformsales %>% 
  replace_na(list(cum_global_plat_sum = 0)) %>%
  filter(cum_global_plat_sum >0) %>% 
  ggplot()+
  geom_col(aes(x = total_other_platform_sum,
             y = fct_reorder(Platform,cum_other_plat_sum),
             fill = factor(Genre),
             group = Platform))+
  scale_fill_manual(values = c("red", "green","blue","black", "yellowgreen", "deeppink","cyan","darkmagenta","steelblue","midnightblue","gold", "green4"))+
  labs(x = "", 
       y = "",
       fill = "Genre",
       caption = "Plot created by D.B. & B.A., Data from Vgsales",
       title ="Other Region Sales by Platform")

p5 <- regionplatformsales %>% 
  filter(total_global_platform_sum >0) %>% 
  ggplot()+
  geom_col(aes(x = total_global_platform_sum,
             y = fct_reorder(Platform,cum_global_plat_sum),
             fill = factor(Genre),
             group = Platform))+
  scale_fill_manual(values = c("red", "green","blue","black", "yellowgreen", "deeppink","cyan","darkmagenta","steelblue","midnightblue","gold", "green4"))+
  labs(x = "", 
       y = "",
       fill = "Genre",
       caption = "Plot created by D.B. & B.A., Data from Vgsales",
       title ="Global Sales by Platform")

p1
p2
p4
p3
p5

```

From these graphs, we see that in North America, the top genres are sold consistently well in all platforms, especially Xbox 360. In Europe and other regions, we see similar patterns with top three genres selling well on all platforms, especially PS3 and PS2 respectively. In Japan, Role playing outsold all the genres in every platform with DS seeing the major part of sales. 

It can be understood that these genres have franchises that bring the majority of the sales so we wanted to take a look at how well the leading franchises of these genres did in terms of global sales. In addition, we looked to understand how our favorite franchises ranked up against each other. 


```{r, fig.alt = "Column chart showing the total global sales of popular video game franchises.", echo = FALSE}
vgsales %>%
  mutate(Name = ifelse(Name == "Minecraft: Story Mode", "Minecraft", Name)) %>%
  mutate(Name = ifelse(Name %in% c("Grand Theft Auto V",
                                   "Grand Theft Auto: Liberty City Stories",
                                   "Grand Theft Auto 2",
                                   "Rockstar Games Double Pack: Grand Theft Auto III & Grand Theft Auto Vice City",
                                   "Grand Theft Auto: Chinatown Wars",
                                   "Grand Theft Auto: San Andreas",
                                   "Grand Theft Auto: Vice City Stories",
                                   "Grand Theft Auto IV","Grand Theft Auto: San Andreas",
                                   "Grand Theft Auto: Mission Pack #1, London 1969",
                                   "Grand Theft Auto: Vice City",
                                   "Grand Theft Auto III"), "Grand Theft Auto", Name)) %>% 
  mutate(Name = ifelse(Name %in% c("Tetris", "Tetris Plus", 
                                   "Tetris DS","Tetris Worlds", 
                                   "Tetris DX", "Tetris 2 (All region sales)",
                                   "Tetris 2", "The Next Tetris", 
                                   "Tetris Party Deluxe", "Tetris Axis",
                                   "Tetris Online", "Tetrisphere",
                                   "Tetris 2 + Bombliss", "The New Tetris",
                                   "Tetris 2 (weekly jp sales)", 
                                   "Tetris Attack", "Tetris Worlds",
                                   "Magical Tetris Challenge", 
                                   "Tetris Worlds (Online Edition)", 
                                   "Star Wars: The Clone Wars & Tetris Worlds",
                                   "Tetris Attack", "Puyo Puyo Tetris",
                                   "Tetris Evolution", "Tetris Ultimate"),
                       "Tetris",Name)) %>% 
  mutate(Name = ifelse(Name %in% c("Super Mario Bros.",
                                   "New Super Mario Bros.",
                                   "New Super Mario Bros. Wii", 
                                   "Super Mario Bros. 3", 
                                   "New Super Mario Bros. 2", 
                                   "Super Mario Bros. 2",
                                   "Super Mario Bros. 3",
                                   "New Super Mario Bros. U",
                                   "Super Mario Bros.: The Lost Levels",
                                   "Mario Bros",
                                   "Classic NES Series: Super Mario Bros.",
                                   "Puzzle & Dragons Z + Super Mario Bros. Edition",
                                   "Famicom Mini: Super Mario Bros. 2",
                                   "Famicom Mini: Mario Bros."),"Mario", Name)) %>%
  mutate(Name = ifelse(Name %in% c("Pokemon Red/Pokemon Blue",
                                   "Pokemon Gold/Pokemon Silver",
                                   "Pokemon Diamond/Pokemon Pearl",
                                   "Pokemon Ruby/Pokemon Sapphire",
                                   "Pokemon Black/Pokemon White",
                                   "Pokemon X/Pokemon Y",
                                   "Pokemon HeartGold/Pokemon SoulSilver",
                                   "Pokemon Omega Ruby/Pokemon Alpha Sapphire",
                                   "Pokemon FireRed/Pokemon LeafGreen",
                                   "Pokemon Black 2/Pokemon White 2",
                                   "Pokemon Stadium",
                                   "Pokemon Pinball",
                                   "Pokemon Mystery Dungeon: Explorers of Time/Explorers of Darkness",
                                   "Pokemon Snap",
                                   "Pokemon Ranger: Shadows of Almia",
                                   "Pokemon Ranger",
                                   "Pokemon Mystery Dungeon: Red/Blue Rescue Team",
                                   "Pokemon Battle Revolution",
                                   "Pokemon Mystery Dungeon: Explorers of Sky",
                                   "Pokemon Pinball: Ruby & Sapphire",
                                   "Pokemon Ranger: Guardian Signs",
                                   "Pokemon Mystery Dungeon: Gates to Infinity",
                                   "Pokemon XD: Gale of Darkness",
                                   "Pokemon Super Mystery Dungeon",
                                   "Pokemon Rumble Blast",
                                   "Pokemon Conquest",
                                   "Pokemon Card GB2: Here Comes Team GR!",
                                   "Pokemon Dash",
                                   "Pokemon Trozei!",
                                   "Pokemon Art Academy",
                                   "Pokemon Puzzle League", 
                                   "Pokemon Channel", 
                                   "Battle & Get! Pokemon Typing DS",
                                   "Pokemon Box: Ruby & Sapphire",
                                   "Pokemon Rumble World",
                                   "Pok?mon Platinum Version",
                                   "Pok?mon Emerald Version",
                                   "Pok?mon Crystal Version",
                                   "Pok?mon Stadium 2","Pok?mon Colosseum",
                                   "Pok?mon Mystery Dungeon: Blue Rescue Team", 
                                   "Pok?Park Wii: Pikachu's Adventure",
                                   "Pok?mon Mystery Dungeon: Red Rescue Team (US weekly sales)",
                                   "Pokken Tournament", 
                                   "PokePark 2: Wonders Beyond",
                                   "Pok?mon I Choose You / Squirtle Squad Game Boy Advance Video", 
                                   "Pok?mon Beach Blank-out Blastoise / Go West Young Meowth Game Boy Advance Video", 
                                   "Pok?mon: Johto Photo Finish: Game Boy Advance Video",
                                   "Pok?mon: For Ho-Oh the Bells Toll!: Game Boy Advance Video", 
                                   "Hey You, Pikachu!"), "Pokemon", Name)) %>% 
  mutate(Name = ifelse(Name %in% c("Pac-Man Collection","Pac-Man: Adventures in Time","Ms. Pac-Man","Pac-Man","Jr. Pac-Man","Pac-Man Player's Choice Bundle","Pac-Man Party","Classic NES Series: Pac-Man","Pac-Man All-Stars","Pac-Man Fever","Pac-Pix","Ms. Pac-Man: Maze Madness / Pac-Man World","Ms. Pac-Man: Maze Madness","Pac-Man & Galaga Dimensions","Pac-Man Party 3D","Pac-Man Pinball Advance","Pac-Man World 2","Pac-Man Power Pack","Pac-Man World 3","Pac-Man and the Ghostly Adventures 2","PAC-MAN and the Ghostly Adventures","Pac-Man World Rally"), "Pac-Man", Name)) %>%
  mutate(Name = ifelse(Name %in% c("Call of Duty: Modern Warfare 3","Call of Duty: Black Ops","Call of Duty: Black Ops 3","Call of Duty: Black Ops II","Call of Duty: Modern Warfare 2","Call of Duty: Ghosts","Call of Duty 4: Modern Warfare","Call of Duty: Advanced Warfare","Call of Duty: World at War","Call of Duty: Finest Hour","Call of Duty 3","Call Of Duty 2: Big Red One","Call of Duty: Roads to Victory","Call of Duty 2","Call of Duty: World at War","Call of Duty: World at War Final Fronts","Call of Duty Black Ops: Declassified","Call of Duty: Modern Warfare: Reflex Edition", "Call of Duty: Modern Warfare: Mobilized","Call of Duty: Modern Warfare 3: Defiance","Call of Duty: The War Collection","Call of Duty: Modern Warfare Trilogy"), "Call of Duty", Name)) %>% 
  mutate(Name = ifelse(Name %in% c("The Sims 3","The Sims: Unleashed","MySims","The Sims: Vacation","The Sims: Livin Large","The Sims 4","The Sims","The Sims: Bustin' Out","The Sims: House Party","The Sims 2: Pets","The Sims: Makin' Magic","The Sims: Hot Date","The Urbz: Sims in the City","The Sims 3: Late Night Expansion Pack","The Sims: Superstar","MySims","MySims Kingdom","The Sims 2","The Sims 2: Nightlife","The Sims 2: Castaway","MySims Agents","The Sims 2: Apartment Pets","The Sims: Medieval","The Sims 3: High-End Loft Stuff","The Sims 3: Generations","MySims Racing","The Sims 3: World Adventures","The Sims 2: Open for Business","The Sims 3: Ambitions","The Sims 3: Outdoor Living Stuff","The Urbz: Sims in the City (all regions sales)","The Sims 3: Town Life Stuff","MySims SkyHeroes","MySims Party","The Sims 4: Get Together","The Sims 3: Seasons","The Sims Medieval: Pirates and Nobles","MySims Collection","The Sims 2: Apartment Life","The Sims 2: Ikea Home Stuff","The Sims 3: Fast Lane Stuff","The Sims 3: Barnacle Bay","The Sims 2: Seasons","The Sims 2: University","The Sims Deluxe","The Sims 4: Spa Day","The Sims 2: Glamour Life Stuff","The Sims 2: Bon Voyage","The Sims: Complete Collection","The Sims 2: Happy Holiday Stuff","The Sims 2: Family Fun Stuff","The Sims 4: Luxury Party Stuff"),"The Sims", Name)) %>% 
  mutate(Name = ifelse(Name %in% c("Final Fantasy VII","Final Fantasy X","Final Fantasy VIII","Final Fantasy XII","Final Fantasy XIII","Final Fantasy IX","Final Fantasy X-2","Final Fantasy III","Crisis Core: Final Fantasy VII","Final Fantasy XIII-2","Final Fantasy V","Final Fantasy Tactics","Dissidia: Final Fantasy","Final Fantasy Tactics Advance","Final Fantasy II","Final Fantasy XIV: A Realm Reborn","Final Fantasy: Crystal Chronicles","Dirge of Cerberus: Final Fantasy VII","Final Fantasy XII: Revenant Wings","Final Fantasy X / X-2 HD Remaster","Final Fantasy IV","Final Fantasy I & II: Dawn of Souls","Final Fantasy Tactics: The War of the Lions","The Final Fantasy Legend","Lightning Returns: Final Fantasy XIII","Final Fantasy Type-0","Dissidia 012: Duodecim Final Fantasy","Final Fantasy","Final Fantasy XI: Wings of the Goddess","Final Fantasy Anthology","Final Fantasy Tactics A2: Grimoire of the Rift","Final Fantasy Legend III","Final Fantasy Crystal Chronicles: Ring of Fates","Final Fantasy VII International","Final Fantasy IV Advance","Final Fantasy: The 4 Heroes of Light","Final Fantasy Anniversary Edition","Final Fantasy IV: The Complete Collection","Final Fantasy Chronicles","Final Fantasy Explorers","Final Fantasy II Anniversary Edition","Theatrhythm: Final Fantasy","Final Fantasy Crystal Chronicles: Echoes of Time","Final Fantasy Collection","Final Fantasy Crystal Chronicles: The Crystal Bearers","Final Fantasy Adventure","Theatrhythm Final Fantasy: Curtain Call","Final Fantasy XI","Final Fantasy IV","Final Fantasy XI: Wings of the Goddess","Lightning Returns: Final Fantasy XIII","Final Fantasy X-2: International + Last Mission","Final Fantasy VI Advance","Final Fantasy X International","Final Fantasy XI: Ultimate Collection","Final Fantasy V Advance","Final Fantasy XI: Chains of Promathia","Final Fantasy Fables: Chocobo Tales","Final Fantasy XI: Treasures of Aht Urhgan","Final Fantasy Fables: Chocobo's Dungeon","Final Fantasy XI","Final Fantasy XI: Vana'diel Collection","Final Fantasy XI: Rise of the Zilart","Final Fantasy XIV: Heavensward","Dragon Quest & Final Fantasy in Itadaki Street Portable","Final Fantasy XII International Zodiac Job System","Final Fantasy VI","Final Fantasy XI: Vana'diel Collection 2008","Dissidia: Final Fantasy Universal Tuning","Final Fantasy XI: All-In-One Pack 2006"), "Final Fantasy", Name)) %>% 
  mutate(Name = ifelse(Name %in% c("FIFA 16","FIFA 14","FIFA 12","FIFA 15","FIFA Soccer 13","FIFA Soccer 11","FIFA 17","FIFA Soccer 06","FIFA Soccer 07","FIFA Soccer 2005","FIFA Soccer 10","FIFA Soccer 2004","FIFA Soccer 2003","FIFA Soccer 08","FIFA Soccer 09","FIFA 2000","FIFA Street","2010 FIFA World Cup South Africa","FIFA Soccer 09 All-Play","FIFA Soccer World Championship","FIFA: Road to World Cup 98","2014 FIFA World Cup Brazil","FIFA Soccer 2002","FIFA Soccer","2002 FIFA World Cup","FIFA: Road to  World Cup 98","FIFA World Cup Germany 2006","FIFA 99","FIFA Soccer 64","FIFA Street 2","FIFA Street 3","FIFA 2001 Major League Soccer","FIFA Soccer 96","FIFA Soccer 97","FIFA 06: Road to FIFA World Cup","FIFA 06 Soccer","Project FIFA World Cup","FIFA 2001: Major League Soccer","FIFA Manager 12","FIFA Soccer 2003 (weekly jp sales)","FIFA World Cup: Germany 2006"), "Fifa", Name)) %>% 
  group_by(Name) %>% 
  summarise(tgvs = sum(Global_Sales)) %>% 
  filter(Name =="Minecraft" | Name == "Grand Theft Auto" | Name == "Tetris"|
           Name == "Mario" | Name == "Pokemon" | Name == "Pac-Man" | 
           Name == "Call of Duty" | Name == "The Sims" | Name == "Final Fantasy" |
           Name == "Fifa") %>% 
  arrange(desc(tgvs)) %>% 
  ggplot(aes(x = tgvs,
             y = fct_reorder(Name, tgvs),
             fill=factor(Name))) +
  geom_col() +
  scale_fill_manual(values = c("gray45","black", "lightblue", "orange", "red","darkgreen","yellow","yellow","purple", "green")) +
  labs(x = "", 
       y = "",
       caption = "Plot created by D.B. & B.A., Data from Vgsales",
       title = "Most Popular Franchises by Global Sales") +
  theme(legend.position = "none")
```

Based on this graph, we saw that the most selling franchise was Call of Duty followed by Pokemon. This makes sense from our personal experience that these franchises would be the most selling. It was interesting to see that a Shooter genre was the highest selling despite Action being the fan favorite. 

After learning this, we wanted to take a look at the most selling individual games and determine which games from these franchises sold the best. 


```{r, fig.width = 8, fig.alt = "Column chart showing the total global sales of the  top ten most popular video games.", echo = FALSE}
vgsales %>% 
  group_by(Name,Genre) %>% 
  summarize(total_gs = sum(Global_Sales)) %>% 
  filter(total_gs>30) %>% 
  arrange(desc(total_gs)) %>% 
  ggplot(aes(x = total_gs,
             y = fct_reorder(Name, total_gs),
             fill = Genre)) +
  geom_col() +
  scale_fill_manual(values = c("red","yellow","deeppink", "cyan",
                              "darkmagenta","steelblue","midnightblue","gold")) +
  labs(x = "", 
       y = "", 
       caption = "Plot created by D.B. & B.A., Data from Vgsales",
       title = "Most Popular Games by Global Sales") 
```

From this, we see that Wii Sports is most selling followed by Grand theft Auto V. Individual games from some of the major franchises are clearly represented with some repeated genres. It can be seen that highest selling games belong to certain genres and that consumers prefer one over another. 

### Lessons learned: 

We learned how to create multi-layered charts that show more than just a x and y axis, where a third aspect of how one variable can affect the outcome of a chart and give more information to a story. Also adding animation to static plots in order to help visualize a better story of data over time. 

### Challenges: 

Throughout our exploration of the data, we saw several limitations that prevented us from doing exactly what we wanted to do. These included the types of variables used for genre as what constituted what as a genre in games as well as multi-genred games. From our experience, we thought that some games could be classified as other genres from what it was in the dataset. The dataset was quite limited as we wanted look at the sales in specific regions around the world but there was only specific details about major contributors to the industry. 


### Conclusion: 

From our explorations of video game sales, we can conclude that the majority of consumers greatly enjoy some genres over others with Action, Sports, and Shooters taking the most global sales. It seems as if individuals find entertainment in doing things that are intangible to reality. As we look at different locations around the world, the popularity of genres, platforms, and individual games are quite different, giving evidence to differing cultural and social values. Furthermore, the advancement of technology towards quality and interface capabilities play a major role in appealing to consumer and video game enthusiast around the world. 

