---
title: "Final Project Graphs"
author: "Brian Anderson and Daniel Beidelshcies"
output: 
  html_document:
    df_print: paged
    code_download: true
    theme: cerulean
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=TRUE, message=FALSE, warning=FALSE)
```

```{r libraries}
library(tidyverse)     # for graphing and data cleaning
library(lubridate)     # for date manipulation
library(ggthemes)      # for even more plotting theme
library(gganimate)     # for adding animation layers to ggplots
library(transformr)    # for "tweening" (gganimate)
library(gifski)        # for creating the gif
library(shiny)         # for creating interactive apps
library(readr)
theme_set(theme_minimal()) # My favorite ggplot() theme :)
```

## Data choice: Video Game Sales

```{r}
vgsales <- read_csv("vgsales.csv", col_types = cols(Year = col_integer()))
vgsales
```

### Final Project Outline

#### Title: 

In the YAML section of the document. A descriptive title & list of all group members.

*Exploring the World of Video Games*

Daniel Beidelschies and Brian Anderson

#### Introduction and Background: 

An introduction that motivates & outlines a clear, specific set of research questions. Also, provide some background on your topic.

##### Introduction

Since the advancement of technology, the creation of video games has played an influential role in being a source of entertainment for everyone around the world. From the multitude of video games that have been created, everyone is able to find something that they enjoy in order to pass time, have fun, and fulfill their creative needs. With this idea, we want to look at how well different games, genres, and platforms did when it came to releasing games for everyone to play. This will allow us to understand how popularity ranks up between different types of games based on the sales around of the world. 

##### Background information

Since our individual discovery, we have always had interest in video games and wondered how the sales of these games differed across the world as we grew up in different regions. 

#### Research Questions

What are the most popular games over the years?

What are the most selling platforms of games over the years across different regions?

What is the most selling genre of games over the years across different regions?

Who is the most selling publisher of games over the years across different regions?

#### Data collection: 

Specification of your data sources and collection process.

Scraping from Kaggle and importing to a dataset

#### Analysis: 

This is the bulk of the report which either has a presentation of the group’s key findings and take-aways or gives the detail of how someone would interact and what people should take away from the shiny app. If you choose to do a shiny app, be sure to include a link to the shinyapps.io site.


```{r, fig.alt = "Animated BAr Chart showin cumulative video game sales by genre from 1980-2020.", echo=FALSE}
vgsales %>%
  group_by( Year,Genre) %>% 
  summarize(total_genre_sum = sum(Global_Sales)) %>% 
  ungroup() %>% 
  complete(Year,Genre) %>% 
  arrange(Genre,Year) %>% 
  group_by(Genre) %>% 
  replace_na(list(total_genre_sum = 0)) %>% 
  mutate(cum_genre_sum = cumsum(total_genre_sum)) %>%
  filter(cum_genre_sum > 0) %>% 
  filter(n()>=13) %>% 
  top_n(n = 12,wt = cum_genre_sum) %>% 
  arrange(Year,desc(cum_genre_sum)) %>% 
  mutate(rank = 1:n()) %>% 
  ggplot(aes(x = cum_genre_sum,
             y = factor(rank),
             fill = Genre,
             group = Genre)) +
  geom_col() +
  geom_text(aes(label = Genre),
            color = "white",
            hjust = "right") +
  labs(title = "Popular Video Game Genres based on Worldwide Sales",
       subtitle = "Year: {frame_time}",
       x = "", 
       y = "",
       caption = "Plot created by Brian Anderson, Data from Vgsales") + 
  transition_time(Year)

```

Genre: Looking at the total sales in North America, Europe, Japan, and other Global regions by Genre

```{r}
regiongenresales <- vgsales %>%
  group_by( Year,Genre) %>% 
  summarize(total_NA_genre_sum = sum(NA_Sales), 
            total_EU_genre_sum = sum(EU_Sales), 
            total_JP_genre_sum = sum(JP_Sales),
            total_other_genre_sum = sum(Other_Sales)) %>% 
  ungroup() %>% 
  complete(Year,Genre) %>% 
  arrange(Genre,Year) %>% 
  group_by(Genre) %>% 
  replace_na(list(total_NA_genre_sum = 0, total_EU_genre_sum  = 0, total_JP_genre_sum  = 0, total_other_genre_sum = 0)) %>% 
  mutate(cum_NA_genre_sum = cumsum(total_NA_genre_sum), cum_EU_genre_sum = cumsum(total_EU_genre_sum),cum_JP_genre_sum = cumsum(total_JP_genre_sum),cum_other_genre_sum = cumsum(total_other_genre_sum)) %>% 
  filter(cum_NA_genre_sum > 0)


g1 <- ggplot(regiongenresales)+
  geom_col(aes(x = cum_NA_genre_sum,
             y = Genre,
             fill = Genre,
             group = Genre))+
  labs(x="", y= "", title ="North American Sales by Genre")+
  theme(legend.position = "none")

g2 <- ggplot(regiongenresales)+
  geom_col( aes(x = cum_EU_genre_sum,
             y = Genre,
             fill = Genre,
             group = Genre))+
  labs(x="", y= "", title ="European Sales by Genre")+
  theme(legend.position = "none")

g3 <- ggplot(regiongenresales)+
  geom_col(aes(x = cum_JP_genre_sum,
             y = Genre,
             fill = Genre,
             group = Genre))+
  labs(x="", y= "", title ="Japanese Sales by Genre")+
  theme(legend.position = "none")

g4 <- ggplot(regiongenresales)+
  geom_col(aes(x = cum_other_genre_sum,
             y = Genre,
             fill = Genre,
             group = Genre))+
  labs(x="", y= "", title ="Other Region Sales by Genre")+
  theme(legend.position = "none")


g1
g2
g3
g4

```

Platform: Looking at the total sales in North America, Europe, Japan, and other Global regions by platform

```{r}
regionplatformsales <- vgsales %>%
  group_by( Year,Platform) %>% 
  summarize(total_NA_platform_sum = sum(NA_Sales), 
            total_EU_platform_sum = sum(EU_Sales), 
            total_JP_platform_sum = sum(JP_Sales),
            total_other_platform_sum = sum(Other_Sales)) %>% 
  ungroup() %>% 
  complete(Year,Platform) %>% 
  arrange(Platform,Year) %>% 
  group_by(Platform) %>% 
  replace_na(list(total_NA_platform_sum = 0, total_EU_platform_sum  = 0, total_JP_platform_sum  = 0, total_other_platform_sum = 0)) %>% 
  mutate(cum_NA_platform_sum = cumsum(total_NA_platform_sum), cum_EU_platform_sum = cumsum(total_EU_platform_sum),cum_JP_platform_sum = cumsum(total_JP_platform_sum),cum_other_platform_sum = cumsum(total_other_platform_sum))


p1 <- ggplot(regionplatformsales)+
  geom_col(aes(x = cum_NA_platform_sum,
             y = Platform,
             fill = Platform,
             group = Platform))+
  labs(x="", y= "", title ="North American Sales by Platform")+
  theme(legend.position = "none")

p2 <- ggplot(regionplatformsales)+
  geom_col( aes(x = cum_EU_platform_sum,
             y = Platform,
             fill = Platform,
             group = Platform))+
  labs(x="", y= "", title ="European Sales by Platform")+
  theme(legend.position = "none")

p3 <- ggplot(regionplatformsales)+
  geom_col(aes(x = cum_JP_platform_sum,
             y = Platform,
             fill = Platform,
             group = Platform))+
  labs(x="", y= "", title ="Japanese Sales by Platform")+
  theme(legend.position = "none")

p4 <- ggplot(regionplatformsales )+
  geom_col(aes(x = cum_other_platform_sum,
             y = Platform,
             fill = Platform,
             group = Platform))+
  labs(x="", y= "", title ="Other Region Sales by Platform")+
  theme(legend.position = "none")


p1
p2
p3
p4

```


```{r, echo=FALSE}
regionpublishersales <- vgsales %>%
  group_by( Year,Publisher) %>% 
  summarize(total_NA_publisher_sum = sum(NA_Sales), 
            total_EU_publisher_sum = sum(EU_Sales), 
            total_JP_publisher_sum = sum(JP_Sales),
            total_other_publisher_sum = sum(Other_Sales)) %>% 
  ungroup() %>% 
  complete(Year,Publisher) %>% 
  arrange(Publisher,Year) %>% 
  group_by(Publisher) %>% 
  replace_na(list(total_NA_publisher_sum = 0, total_EU_publisher_sum  = 0, total_JP_publisher_sum  = 0, total_other_publisher_sum = 0)) %>% 
  mutate(cum_NA_publisher_sum = cumsum(total_NA_publisher_sum), cum_EU_publisher_sum = cumsum(total_EU_publisher_sum),cum_JP_publisher_sum = cumsum(total_JP_publisher_sum),cum_other_publisher_sum = cumsum(total_other_publisher_sum))  


pu1 <- ggplot(regionpublishersales)+
  geom_col(aes(x = cum_NA_publisher_sum,
             y = Publisher,
             fill = Publisher,
             group = Publisher))+
  labs(x="", y= "", title ="North American Sales by Publisher")+
  theme(legend.position = "none")

pu2 <- ggplot(regionpublishersales)+
  geom_col( aes(x = cum_EU_publisher_sum,
             y = Publisher,
             fill = Publisher,
             group = Publisher))+
  labs(x="", y= "", title ="European Sales by Publisher")+
  theme(legend.position = "none")

pu3 <- ggplot(regionpublishersales)+
  geom_col(aes(x = cum_JP_publisher_sum,
             y = Publisher,
             fill = Publisher,
             group = Publisher))+
  labs(x="", y= "", title ="Japanese Sales by Publisher")+
  theme(legend.position = "none")

pu4 <- ggplot(regionpublishersales)+
  geom_col(aes(x = cum_other_publisher_sum,
             y = Publisher,
             fill = Publisher,
             group = Publisher))+
  labs(x="", y= "", title ="Other Region Sales by Publisher")+
  theme(legend.position = "none")

```

Individual Games: Looking at what games were the most popular and had the most sales over time

```{r}
vgsales %>% 
  filter(Global_Sales > 28) %>% 
  ggplot(aes(x = Global_Sales,
             y = Name)) +
  geom_col() +
  labs(title = "Top Most Popular Games from 1980-2020",
       x = "Global Sales (Millions)",
       y = "")
```

